---
title: "Data Visualization"
author: "Amritpal Sandhu"
date: "2022-08-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r}
library(tidyverse)
library(here)
library(janitor)

library(patchwork)
library(ggrepel)
library(gghighlight)
library(paletteer)
library(ggExtra)
library(ggbeeswarm)

library(gapminder)
library(sf)
```

```{r}
lizards <- read_csv(here::here("data_tidy", "lizards.csv"))
```

```{r}
ggplot(data = lizards) + geom_point(aes(x = total_length, y = weight))
```

```{r}
# Create a histogram of total lengths for all lizards in the dataset.

# Put variables inside of aes not things that refer to constants 

plot_1 <- ggplot(data = lizards, aes(x = total_length)) + 
  geom_histogram(color = "orange",
                 fill = "purple",
                 size = .3,
                 linetype = "dotted")

plot_1
```

```{r}
ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point(shape = 22, 
             color = "cyan4",
             fill = "yellow",
             size = 4,
             alpha = .4)
```

```{r}
ggplot(data = lizards, aes(x = total_length, 
                           y = weight)) + 
  geom_point(aes(color = common_name, 
                 size = total_length),
             alpha = 0.5) +
  theme_minimal()
```

# Faceting and ~ means a function of

```{r}
ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point() +
  facet_wrap(~common_name, scales = "free")
```

# Re-ordering

```{r}
lizard_count <- lizards |>
  group_by(common_name) |>
summarize(
  counts = n()
)
 # OR

lizard_count_2 <- lizards |>
  count(common_name, site, tail)
```

```{r}
## Make column graph

ggplot(data = lizard_count_2, 
       aes(y = fct_reorder(common_name, n),
                                x = n)) +
  geom_col()

## Factor(fct) to reorder things based on values
```

Create a scatterplot, faceted by species, for total length versus weight. The color and shape  of points should differ based on the species (common name). Change the theme. Make each faceted panel have flexible axis limits (scales should be free).

```{r}
ggplot(data = lizards, 
       aes(x = total_length,
           y = weight)) + 
  geom_point(aes(color = common_name,
                 shape = ,
                 alpha = 0.6,
                 size = 3)) +
  facet_wrap(~common_name, scales = "free") + theme_light() + 
  labs(x = "Total Length (mm)",
       y = "Weight (grams)", 
       title = "Lizard sizes", 
       subtitle = "All lizards follow the standard length-weight relationship",
       caption = "Data : Jordana ...")
  
```

```{r}
### Fun with scales
ggplot(data = lizards, 
       aes(x = total_length,
           y = weight)) + 
  geom_point(aes(color = weight)) +
  scale_color_gradient(low = "purple", high = "yellow")

## Two color gradient
ggplot(data = lizards, 
       aes(x = total_length,
           y = weight)) + 
  geom_point(aes(color = weight)) +
  scale_color_gradientn(colors = c("magenta", "cyan4", "dodgerblue", "green", "yellow", "darkorchid3"))

## Step
ggplot(data = lizards, 
       aes(x = total_length,
           y = weight)) + 
  geom_point(aes(color = weight)) +
  scale_color_steps(low = "red", high = "black")

ggplot(data = lizards, 
       aes(x = total_length,
           y = weight)) + 
  geom_point(aes(color = total_length)) +
  scale_color_steps2(low = "purple",
                     mid = "white",
                     high = "orange",
                     midpoint = 150, 
                     breaks = c(50, 75, 150, 180, 220, 280))


ggplot(data = lizards, 
       aes(x = total_length,
           y = weight)) + 
  geom_point(aes(color = total_length)) +
  scale_color_stepsn(colors = c("orange", "red", "purple"),
                     breaks = c(50, 55, 80, 140, 200))
## the breaks could also be a sequence witten out: breaks = seq(from = 0, to = 300, by = 60)
```

## Make a horizontal boxplot with common name on the y- axis, total length on the x-axis, with boxplot fill changing based on the common name
```{r}
# reorder based on median value with factor reorder
# creating new column that overrides an existing column, reordering factor, and put them in order of median

lizards_median <- lizards |>
  mutate(common_name = fct_reorder(common_name, total_length, .fun = median))

class(lizards_median$common_name)
levels(lizards_median$common_name)

ggplot(data = lizards, aes(x = total_length, 
                           y = common_name)) + geom_boxplot(aes(fill = common_name)) 


ggplot(data = lizards_median, aes(y = common_name, 
                                  x = total_length)) +
  geom_boxplot(aes(fill = common_name), show.legend = FALSE) +
  scale_fill_paletteer_d(palette = "beyonce::X108") +
  theme_minimal()

# View(palettes_d_names)
# Also scale_fill_paletteer_d(palette = "ggprism::")
```

```{r}
ggplot(data = lizards, 
       aes(x = total_length,
           y = weight)) + 
  geom_point(aes(color = weight)) +
  theme(panel.grid = element_line(color = "red"))

# Grid lines color change only at certain points
ggplot(data = lizards, 
       aes(x = total_length,
           y = weight)) + 
  geom_point(aes(color = weight)) +
  theme(panel.grid.major.x = element_line(color = "red"),
        panel.grid.minor.y = element_line(color = "green"), 
        axis.title.x = element_text(color = "purple", size = 10),
        axis.text.y = element_text(color = "green"),
        text = element_text(size = 18),
        panel.background = element_rect(color = "purple", fill = "yellow")) + annotate("text", x = 300, y = 50, label = "Woooooo", color = "blue") +
  geom_vline(xintercept = 250, linetype = "dashed", color = "orange", size = 4)

```

## Repulsive labels with `ggrepel`

```{r}

wws_lizards <- lizards |>
  filter(common_name == "western whiptail",
         site == "sand")

ggplot(data = wws_lizards, aes(y = weight, 
                                  x = total_length)) +
  geom_point() +
  geom_text_repel(aes(label = toe_num))

gapminder |>
  filter(year == 2002, continent == "Europe") |>
  ggplot(aes(x = gdpPercap, y = lifeExp)) +
  geom_point() +
  geom_text_repel(aes(label = country), size = 3)
```



```{r}
ggplot(data = lizards, aes(y = weight, 
                                  x = total_length)) +
  geom_point() +
  gghighlight(toe_num == 250, label_key = toe_num)

# Try this: gghighlight(weight > 30, label_key = toe_num)
```
## Make a map

# can't use read function to read in the data

```{r}
jornada_veg <- read_sf(here("data_raw", "spatial_vegetation", "doc.kml"))

ggplot() + 
  geom_sf(data = jornada_veg, 
          aes(fill = Name),
          color = NA) + 
  theme_minimal() +
  scale_fill_paletteer_d(palette = "ggthemes::manyeys")
  
```

